---
title: "Code to generate the pre-computed objects from the _HumanDEU_ package"
author: "Alejandro Reyes"
date: "2017"
package: "`r BiocStyle::pkg_ver('HumanDEU')`"
abstract: >
  This data package provides pre-computed data objects with relative
  exon usage coefficients and relative spliced-in coefficients. The
  reason for this is that the computation of these coefficients
  is very expensive computationally. We used computational cluster to
  generate these coefficients in a reasonable amount of time.  This
  vignette provides the code used to generate these coefficients.
  However, this vignette only runs the code for the exonic
  regions of the first gene.
output:
  BiocStyle::html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{HumanDEU 1}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r echo=FALSE, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE,
                      cache = FALSE,
                      dev = "png",
                      message = FALSE,
                      error = FALSE,
                      warning = TRUE)
```


# Estimating REUCs

Below are the steps that were followed to estimate relative
exon usage coefficients (REUCs). We load the data objects 
and extract the indexes of the exonic regions of the first 
gene. 

```{r reucs}

library( HumanDEU )
library( DEXSeq )
data( "dxdObjects" )
dxd <- dxd1
gene1Index <-
    which( mcols( dxd )$groupID %in% "ENSG00000000003" )

```

We fit the model for each exonic region using a large prior 
for shrinkage and vector of fixed dispersion 
estimates. We also use a single core for the calculations,
since we are estimating these for a single gene.  

```{r}

dispThis <- rep( 0.1, length.out = length( gene1Index ) )
dispOthers <- rep( 0.1, length.out = length( gene1Index ) )
names( dispThis ) <- rownames( dxd )[gene1Index]
names( dispOthers ) <- rownames( dxd )[gene1Index]
testableVector <- rep( TRUE, length( gene1Index ) )
names( testableVector ) <- rownames( dxd )[gene1Index]
bjp <- SerialParam()

allCoefsWeakShrinkage <-
    fitAllExonsParallel(
        dxd[gene1Index,],
        dispThis,
        dispOthers,
        priorsd=3,
        bjp )

```

We used these coefficients with weak shrinkage
to get dispersion estimates for each exonic region.

```{r disps}

rawDisps <-
    estimateDispersionsParallel(
        dxd[gene1Index,],
        allCoefsWeakShrinkage,
        bjp)

head( rawDisps )

```

Using these dispersion estimates, we run now the fit
with a smaller prior.

```{r reucs2}

dispThis <- pmax( pmin( rawDisps[,"dispThis"], 10 ), 1e-6 )
dispOthers <- pmax( pmin( rawDisps[,"dispOthers"], 10 ), 1e-6 )

allCoefsGoodShrinkage <- fitAllExonsParallel(
    dxd[gene1Index,],
    dispThis,
    dispOthers,
    priorsd=0.05, bjp )

allCoefsSub <-
    HumanDEU:::arrangeInto3DArray(
        dxd[gene1Index,],
        allCoefsGoodShrinkage )

```

We verify that the estimated REUCs are the same
as in the REUCs in the pre-computed object.

```{r veryfi}
data("crossCoefs1")
all( allCoefsSub == crossCoefs1[gene1Index,,] )
```

# Estimating p-values

As for the REUCs and RSICs, the code below demonstrates the functions
that were used to test for differential exon usage. Due to long
running times, we demonstrate the p-value calculation for the exonic
regions of the first gene.

```{r}

dxd <- dxd1

formulaFull <- ~ sample + exon + sex:exon + individual:exon + tissue:exon
formulaNull <- ~ sample + exon + sex:exon + individual:exon

disps <- estimateTestDispersionsParallel(
    dxd[gene1Index,],
    formula=formulaFull,
    bjp )

disps <- data.frame(
    dispThis=disps,
    dispOthers=disps )
rownames(disps) <- gsub("\\d\\.", "", rownames(disps))

pvals <- testForDEUParallel(
    dxd[gene1Index,],
    formulaNull,
    formulaFull,
    disps,
    testableVector,
    bjp )

data("pvals1")

all( pvalsTissues1[gene1Index] == pvals )

```

